Embodied Neuromorphic Benchmarks

### 2015-08-13 paper/paper.tex
Added Chris' edits

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Stewart {et~al.}} %use et al only if is more than 1 author
\def\Authors{Terrence C. Stewart\,$^{1,*}$, Ashley Kleinhans\,$^{2}$ and Chris Eliasmith\,$^1$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Centre for Theoretical Neuroscience, University of Waterloo,
    Waterloo, ON, Canada
    \\
$^{2}$Mobile Intelligent Autonomous Systems group, 
      Council for Scientific and Industrial Research,
      Pretoria, South Africa}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Terrence C. Stewart}
\def\corrAddress{Centre for Theoretical Neuroscience, University of Waterloo,
    Waterloo, ON, Canada}
\def\corrEmail{tcstewar@uwaterloo.ca}

@@

    Evaluating the effectiveness and performance of neuromorphic hardware is
    difficult.  It is even more difficult when the task of interest is an
    embodied task; that is, a task where the output from the neuromorphic
    hardware affects its future input.  However, these embodied situations
    are one of the primary potential uses of neuromorphic hardware.  To address
    this, we present a methodology for embodied benchmarking and a particular
    example of a benchmark for adaptive control of an arbitrary system.  These
    benchmarks are flexible, in that they allow researchers to explicitly
    modify the benchmark to identify particular task domains where particular
    hardware excels.  Furthermore, the benchmarks make use of a hybrid of
    real physical embodiment and a type of ``minimal'' simulation that has been
    shown to lead to robust real-world performance, allowing the same benchmark
    to be used by many researchers.

@@

Neuromorphic hardware holds great promise for a wide variety of applications.
The combination of massively parallel computation and low power consumption
means that there is the potential to have complex algorithms running in
embedded processing situations, without being a significant drain on battery
life.  The question, of course, is to identify what sort of always-on or
interactive functionality is feasible with these devices.

To evaluate applications of this hardware, we need benchmark tasks.  These
tasks must allow us to compare across different instances of neuromorphic
hardware (and potentially across different algorithms implemented in said
hardware).  Having such a metric across instances and algorithms will allow us
to quantitatively compare systems, letting
researchers both measure the progress in the field and also directly
compare competing approaches.

In this paper, we focus on the development of \emph{embodied} benchmarks.
These are dynamic tasks where the output of the neuromorphic hardware \emph{influences
its own future input}.  This is in contrast to standard categorization or pattern
identification tasks, where the input is some fixed sequence, and the main
question is whether the hardware produces the correct output for each input
(or input pattern).

We believe embodied benchmarks should be of particular interest to
neuromorphic research, given that many of the applications of neuromorphic
hardware are in exactly this domain of embedded and interactive control
of robotic or other physical systems.  However, embodiment raises a number
of issues that complicate the development of such benchmarks.  Rather than
simply providing a data file of inputs and desired outputs, the benchmark
must either specify a full physical system for that embodiment, or it must 
provide software for a simulation of that system.  As we discuss below, either
approach is problematic, and addressing this difficulty is the primary goal of
this paper.

@@

An embodied benchmark task is one where the system we are studying has a 
two-way interaction with some sort of environment.  That is, the outputs from
the neuromorphic hardware are sent to the environment where they cause some
sort of effect, the results of which change the subsequent input.  For example,
the outputs might control the movement of a robot, which in turn affects the
sensory data received by the robot.

@@

To define such a benchmark, we need to be explicit about the embodiment.  If
it is a situation where a robot is to be controlled, we need to be explicit
about all the details of that robot.  What motors are there?  How are they
configured?  How strong are they?  What sensors are there?  Where are they
placed?  How accurate are they?  However, even if these questions are
answered, there is a fundamental problem in that \emph{other researchers
need access to that exact robot}.  If a benchmark is to be widely used, other
researchers developing their own hardware should be able to do their own 
testing.

Furthermore, using a physical robot also imposes significant practical
difficulties when performing extensive benchmark testing.  When testing, we
often want to run the same task over and over again, both for robustness and
to see the effects of varying parameters.  With a physical robot, this means
manually setting up the task, letting the test run, gathering the resulting
data, and then resetting the robot back to the initial state.  This means that
issues like battery life become problematic, and not just because there is a
limited amount of time available for testing.  As the battery level changes,
the performance of the robot itself can also change.  Futhermore, for any
rigorous testing of the benchmark, we will want to examine situations where
the system fails.  This means that some of the testing will involve parameter
settings that lead to poor behaviour, which could involve physical damage to
the robot itself.

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should anyone have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
very widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing.  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap of failing to generalize to real situations.

Furthermore, neuromorphic hardware has another constraint that severely limits
the possibilities of simulation.  The normal approach when a simulation is
too simple to reflect reality is to add details to the simulation itself.
Incredibly finely detailed simulations can be created, filling in all of the
details needed.  However, accurate modelling of physical systems can very quickly
become \emph{impractical to run in real time}.
This is a fundamental problem, in that most neuromorphic
hardware is about real-time interactions, and there is no way to slow down
the hardware to match the simulated environment.  This means that even if
we spent the considerable amount of research effort needed to define a
simulated environment for an embodied benchmark, running that simulation 
fast enough to interact with most of the desired hardware would require more
resources than are available.

@@

The above considerations seem to indicate that even though using real-world
physical hardware for benchmarking is problematic, it is still better than
using simplistic simulations which may not generalize to real tasks.  However,
we believe that the solution lies in an alternate approach taken from the
field of evolutionary robotics, known as \emph{Minimal Simulation}
(\citealt{Jakobi97evolutionaryrobotics}).

First, we note that the problem faced here is remarkably similar to a problem
faced by the evolutionary robotics community.  In evolutionary robotics, the
goal is to use genetic algorithms to \emph{evolve} systems that can control
robots to perform various tasks.  These tasks can be as simple as navigation
and obstacle avoidance, but have also included walking,
object identification, and visual tracking [refs???].

However, performing this evolution on real physical robots is problematic
for the same reasons that benchmarks on physical robots are problematic.  The
robots must be reset to the same state each time; they often involve
behaviour that can physically damage the robots; and they take a very long
time to run.  For this reason, attempts were made to evolve using simulated
robots.  However, the general finding was that algorithms that worked on the
simulated robots would not work when run on the real physical robots.  If
the simulations were improved, adding complex physical detail, then it was
possible to generalize to real behaviour; unfortunately, such complex
simulations would run slower than real-time [refs].

To address this problem, \cite{Jakobi97evolutionaryrobotics} proposed the creation of
``minimal'' simulations.  These are simulations where there is variability
\emph{within the simulation itself}.  In other words, we make \emph{poor}
simulations, but ensure that the way in which it is poor is itself
variable.  We then make sure that the controllers work across that whole
range of variability.  ``Instead of trying to eliminate the differences between
simulation and reality, they are acknowledged, and mechanisms are put in place
to prevent evolving controllers from relying on them.'' [ref: jacobi thesis].

With this approach, it was possible to build minimal simulations that would
run faster than real-time and yet also be complex enough that if a system
could successfully control the simulation, it was also likely to successfully
control a real robot.  To achieve this, the simulations are made to be
unreliable in almost every respect.  For example, for a simulation of a simple
motor it would still be the case that if power is applied it would generally
try to spin, but the exact amount of torque, the amount of sensory noise,
the amount of time needed, the amount of static and dynamic friction, and so
on would all be randomly chosen.  A successful controller would have to deal
with this wide range of variability, and if it could handle that variability
then we would have reason to believe it could also handle the real system.

It is also worth noting that a minimal simulation only has to be a
good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
corridor following behaviour, the dynamics of the simulation might
differ wildly from those of reality if the controller hits a wall or
goes round in circles, but this does
not matter, since the controllers we are interested in transferring
across the reality gap will neither
hit walls nor go round in circles.''  If the controller is poor, we do not 
need the simulation to be at all accurate in exactly \emph{how} that poor
behaviour is manifest.  We do not need an exact detailed physics model of
the collision between a robot and a wall, or a detailed model of what happens
to a robot arm when it starts oscillating wildly due to a poor control signal.
All we need is for the simulation to be just good enough to indicate that
things have gone wrong, and thus give a low score to that controller.  This
means that, for example, in a minimal simulation of an eight-legged walking robot,
it is not necessary to have a physics simulation that correctly models what
happens when two legs collide with each other.  Rather, if legs collide
with each other, that is an indication that the walking behaviour is very poor,
so as long as that result is indicated we can greatly simplify the simulation
by not including all the details necessary to deal with these sorts of physical
interactions.

>>>>

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Stewart {et~al.}} %use et al only if is more than 1 author
\def\Authors{Terrence C. Stewart\,$^{1,*}$, Travis DeWolf\,$^{1}$, Ashley Kleinhans\,$^{2}$ and Chris Eliasmith\,$^1$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Centre for Theoretical Neuroscience, University of Waterloo,
    Waterloo, ON, Canada
    \\
$^{2}$Mobile Intelligent Autonomous Systems group, 
      Council for Scientific and Industrial Research,
      Pretoria, South Africa}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Terrence C. Stewart}
\def\corrAddress{Centre for Theoretical Neuroscience, University of Waterloo,
    Waterloo, ON, Canada}
\def\corrEmail{tcstewar@uwaterloo.ca}

@@

    Evaluating the effectiveness and performance of neuromorphic hardware is
    difficult.  It is even more difficult when the task of interest is an
    embodied task; that is, a task where the output from the neuromorphic
    hardware affects its future input through some environment.  However, embodied situations
    are one of the primary potential uses of neuromorphic hardware.  To address
    this, we present a methodology for embodied benchmarking that makes use
    of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
    Minimal simulation has been shown to lead to robust real-world performance,
    while still maintaining the practical advantages of simulation, such as
    making it easy for the same benchmark to be used by many researchers.
    These benchmarks are flexible, in that they allow researchers to explicitly
    modify the benchmark to identify particular task domains where particular
    hardware excels.  To demonstrate the method, we present a novel benchmark
    where the task is to perform motor control on an arbitrary unknown system.

@@

Neuromorphic hardware holds great promise for a wide variety of applications.
The combination of massively parallel computation and low power consumption
means that there is the potential to have complex algorithms running in
embedded processing situations, without being a significant drain on available
energy.  A crucial challenge is to identify what sort of always-on or
interactive functionality can best exploit these devices.

To evaluate applications of neuromorphic hardware, we need benchmark tasks.  These
tasks must allow us to compare across different instances of neuromorphic
hardware (and potentially across different algorithms implemented in that
hardware).  Good benchmarks will allow us to quantitatively compare systems, letting
researchers both measure the progress in the field, and also directly
compare competing approaches.

In this paper, we focus on the development of \emph{embodied} benchmarks.
These are dynamic tasks where the output of the neuromorphic hardware \emph{influences
its own future input} through some environment.  This is in contrast to standard categorization or pattern
identification tasks, where the input is some fixed sequence and the hardware
must produce the correct output for each input (or input pattern).

We believe embodied benchmarks should be of particular interest to
neuromorphic research, given that the most compelling applications of neuromorphic
hardware are likely to be in this domain of embedded and interactive control
of robotic or other physical systems.  However, embodiment itself raises a number
of issues that complicate the development of such benchmarks.  Rather than
simply providing a data file of inputs and desired outputs, the benchmark
must either specify a full physical system for that embodiment, or it must 
provide software for a simulation of that system.  As we discuss below, both
approaches are problematic.  
Describing a method for overcoming these shortcomings is the primary goal of
this paper.

@@

An embodied benchmark task is one where the system we are studying has a 
two-way interaction with some sort of environment.  That is, the outputs from
the neuromorphic hardware are sent to the environment where they cause
an effect, the results of which change the subsequent input.  For example,
the outputs might control the movement of a robot, which in turn affects the
sensory data received by the robot.

@@

To define such a benchmark, we need to be explicit about the embodiment.  If
a robot is to be controlled, we need to be explicit
about all of the details of that robot.  What motors does it have?  How are they
configured?  How strong are they?  What sensors are there?  Where are they
placed?  How accurate are they?  However, even if these questions are
answered, there is a fundamental problem in that \emph{other researchers
need access to that exact robot}.  If a benchmark is to be widely used, other
researchers developing their own neuromorphic hardware should be able to do their own 
testing on the same benchmark system.

Furthermore, using a physical robot imposes significant practical
difficulties when performing extensive benchmark testing.  When testing, we
often want to run the same task over and over again, both for robustness and
to see the effects of varying parameters.  With a physical robot, this means
manually setting up the task, letting the test run, gathering the resulting
data, and then resetting the robot back to the initial state.  Consequently,
issues like battery life become problematic, and not just because there is a
limited amount of time available for testing.  As the battery level changes,
the performance of the robot itself can also change.  Futhermore, for any
rigorous testing of the benchmark, we will want to examine situations where
the system fails.  This means that some of the testing will involve parameter
settings that lead to poor behaviour, which might have the undesireable 
result of causing physical damage.

However, \emph{not} using a real physical embodiment for testing is also
problematic.  First and foremost, without an actual real-world task, why
should we have any confidence that the performance on the benchmark is
reflective of the actual usefulness of the neuromorphic hardware?  It is
widely known that simulations of robots (or other physical systems) are
often \emph{much} easier to control and better-behaved than the real thing [TODO: ref??].  The
field of robotics is filled with algorithms that work well ``in theory,'' but
fail when run on actual hardware.  We do not want a benchmark that falls into
this trap, giving high scores to hardware that does not turn out to
functional well when generalized to real situations.

Furthermore, neuromorphic hardware has another constraint that severely limits
the utility of simulation.  Typically, when a simulation is
too simple to reflect reality, we add details to the simulation itself.
Incredibly finely detailed simulations can be created, filling in all of the
details needed.  However, accurate modelling of physical systems can very quickly
become \emph{impractical to run in real time}.
This is a fundamental problem, in that neuromorphic
hardware is often tied to real-time interactions, and there is no way to slow down
the hardware to match the simulated environment.  This means that even if
we spent the considerable amount of research effort needed to define a
simulated environment for an embodied benchmark, running that simulation 
fast enough to interact with the desired hardware would require more
resources than are available.

@@

The above considerations seem to indicate that even though using real-world
physical hardware for benchmarking is problematic, it is still better than
using simplistic simulations which may not generalize to real tasks.  However,
we believe neuromorphic benchmarking can effectively exploit an approach 
known as \emph{Minimal Simulation}
(\citealt{Jakobi97evolutionaryrobotics}).

This approach was first suggested in the context of evolutionary robotics.  
Notably, the problem faced by embodied neuromorphic benchmarking is remarkably 
similar to that faced earlier by these researchers. 
In evolutionary robotics, the
goal is to use genetic algorithms to \emph{evolve} systems that can control
robots to perform various tasks.  These tasks can be as simple as navigation
and obstacle avoidance, but have also included more difficult tasks such as walking,
object identification, and visual tracking [refs???].

However, performing evolution on real physical robots is problematic
for the same reasons that benchmarks on physical robots are problematic.  The
robots must be reset to the same state each time; they often involve
behaviour that can physically damage the robots; and they take a very long
time to run.  For this reason, attempts were made to evolve algorithms using simulated
robots.  However, the general finding was that algorithms that worked on the
simulated robots would not work when run on the real physical robots.  If
the simulations were improved, adding complex physical detail, then it was
possible to generalize to real behaviour; unfortunately, such complex
simulations would run slower than real-time [refs].

To address this problem, \cite{Jakobi97evolutionaryrobotics} proposed the creation of
``minimal'' simulations.  These are simulations where there is variability
\emph{within the simulation itself}.  In other words, we make \emph{poor}
simulations, but ensure that the way in which they are poor is itself
variable.  We are then in a position to ensure that the controllers work across that whole
range of variability.  ``Instead of trying to eliminate the differences between
simulation and reality, they are acknowledged, and mechanisms are put in place
to prevent evolving controllers from relying on them.'' [ref: jacobi thesis].

With this approach, it became possible to build minimal simulations that would
run faster than real-time and yet also be complex enough that if a system
could successfully control the simulation, it was also likely to successfully
control a real robot.  To achieve this kind of transfer, the simulations were made to be
unreliable in almost every respect.  For example, for a simulation of a simple
motor it would still be the case that if power is applied it would generally
try to spin, but the exact amount of torque, the amount of sensory noise,
the amount of time needed, the amount of static and dynamic friction, and so
on would all be randomly chosen.  A successful controller would have to deal
with this wide range of variability, and if it could handle that variability
then there would be reason to believe it could also handle the real physical system.

It is worth noting that a minimal simulation only has to be a
good simulation \emph{for successful behaviour}.  That is, ``if we are evolving 
corridor following behaviour, the dynamics of the simulation might
differ wildly from those of reality if the controller hits a wall or
goes round in circles, but this does
not matter, since the controllers we are interested in transferring
across the reality gap will neither
hit walls nor go round in circles.''  [ref: jacobi thesis]  If the controller is poor, we do not 
need the simulation to be at all accurate in exactly \emph{how} that poor
behaviour is manifest.  We do not need an exact detailed physics model of
the collision between a robot and a wall, or a detailed model of what happens
to a robot arm when it starts oscillating wildly due to a poor control signal.
All we need is for the simulation to be just good enough to indicate that
things have gone wrong, and thus give a low score to that controller.  This
means that, for example, in a minimal simulation of an eight-legged walking robot,
it is not necessary to have a physics simulation that correctly models what
happens when two legs collide with each other.  Rather, if legs collide
with each other, that is an indication that the walking behaviour is very poor.
As long as that result is indicated we can greatly simplify the simulation
by not including all the details necessary to deal with these sorts of physical
interactions.
